---
title: "Transfer Entrop Flow Analysis"
author: "Mariel Jones"
date: "2024-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
fig_save = "./Figures/"
load_path = "./mef-data/"
```

```{r}
library(astsa)
library(dplyr)
library(xts)
library(lubridate)
library(RTransferEntropy)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(forecastML)
library(cowplot)
library(reshape2)
library(zoo)
library(infotheo)
library(stringr)
library(latex2exp)
library(Hmisc)
library(plotly)
library(multcompView)
```

# Data Import

Including: 10 minute soil moisture, seasonal soil moisture, precipitation, streamflow, bog well WTE, temperature, and snow depth from GR

```{r}
#Import Soil Moisture and PDSI data
## 10 minute soil moisture
smURL  <- "https://pasta.lternet.edu/package/data/eml/edi/989/3/e25d4f15276a3158af412758502a59b0" 
smInfile <- tempfile()
try(download.file(smURL,smInfile,method="curl"))
if (is.na(file.size(smInfile))) download.file(smURL,smInfile,method="auto")

sm <-read.csv(smInfile, header=F, skip=1, sep=",", 
              col.names=c("TIMESTAMP", "S2S_UP_SH", "S2S_UP_DP", "S2S_MI_SH", 
                          "S2S_MI_DP", "S2S_LO_SH", "S2S_LO_DP", "S2N_UP_SH",     
                          "S2N_UP_DP", "S2N_MI_SH", "S2N_MI_DP", "S2N_LO_SH",     
                          "S2N_LO_DP"), check.names=TRUE)
               
unlink(smInfile)
tmp2TIMESTAMP<-as.POSIXct(sm$TIMESTAMP,format="%Y-%m-%d %H:%M:%S")
# Keep the new dates only if they all converted correctly
if(nrow(sm[sm$TIMESTAMP != "",]) == length(tmp2TIMESTAMP[!is.na(tmp2TIMESTAMP)])){sm$TIMESTAMP <- tmp2TIMESTAMP } else {print("Date conversion failed for sm$TIMESTAMP. Please inspect the data and do the date conversion yourself.")} 
#Aggregate to daily
sm$day <- floor_date(sm$TIMESTAMP, "day")
sm_daily = data.frame(sm) %>%
  group_by(day) %>%
  summarise_all(mean) %>%
  select(-TIMESTAMP)


#Seasonal Soil Moisture (previously processed into plant available water)
seasm <-read.csv('./mef-data/seasonal/mef_seasonal_soil_moisture_processed.csv', header=T, sep=",")
         
# Precipitation
precipURL  <- "https://pasta.lternet.edu/package/data/eml/edi/563/6/f3a58ff544a4ddd3475d265da61bf40e" 
precipInfile <- tempfile()
try(download.file(precipURL,precipInfile,method="curl"))
if (is.na(file.size(precipInfile))) download.file(precipURL,precipInfile,method="auto")

precip <-read.csv(precipInfile , header=F, skip=1, sep = ",", 
               col.names=c("DATE", "NADP_PCP", "South_PCP", "North_PCP",     
                    "NADP_Flag", "South_Flag", "North_Flag") ,check.names=TRUE)
unlink(precipInfile)
tmp3DATE<-as.POSIXct(precip$DATE,format="%Y-%m-%d")
# Keep the new dates only if they all converted correctly
if(nrow(precip[precip$DATE != "",]) == length(tmp3DATE[!is.na(tmp3DATE)])){precip$DATE <- tmp3DATE } else {print("Date conversion failed for precip$TIMESTAMP. Please inspect the data and do the date conversion yourself.")}                                 
if (class(precip$South_PCP)=="factor") precip$South_PCP <-as.numeric(levels(precip$South_PCP))[as.integer(precip$South_PCP) ]               
if (class(precip$South_PCP)=="character") precip$South_PCP <-as.numeric(precip$South_PCP)
if (class(precip$South_Flag)!="factor") precip$South_Flag<- as.factor(precip$South_Flag)
                
# Convert Missing Values to NA for non-dates
precip$South_PCP <- ifelse((trimws(as.character(precip$South_PCP))==trimws("NA")),NA,precip$South_PCP)         

# Streamflow
stream <-read.csv('./mef-data/daily/Streamflow_daily.csv', header=T, sep=",")
stream = stream %>%
  filter(Watershed == 'S2')

# Bog well WTE
bogURL  <- "https://pasta.lternet.edu/package/data/eml/edi/562/3/671f15337a677da71852de506a8d9b05" 
bogInfile <- tempfile()
try(download.file(bogURL,bogInfile,method="curl"))
if (is.na(file.size(bogInfile))) download.file(bogURL,bogInfile,method="auto")

bog <-read.csv(bogInfile, header=F,skip=1, sep=",",
               col.names=c("PEATLAND", "DATE", "WTE", "FLAG"), check.names=TRUE)
               
unlink(bogInfile)
if (class(bog$PEATLAND)!="factor") bog$PEATLAND<- as.factor(bog$PEATLAND)                                   
tmp1DATE<-as.Date(bog$DATE,format="%Y-%m-%d")
# Keep the new dates only if they all converted correctly
if(nrow(bog[bog$DATE != "",]) == length(tmp1DATE[!is.na(tmp1DATE)])){bog$DATE <- tmp1DATE } else {print("Date conversion failed for bog$DATE. Please inspect the data and do the date conversion yourself.")}
if (class(bog$WTE)=="factor") bog$WTE <-as.numeric(levels(bog$WTE))[as.integer(bog$WTE) ]               
if (class(bog$WTE)=="character") bog$WTE <-as.numeric(bog$WTE)
if (class(bog$FLAG)!="factor") bog$FLAG<- as.factor(bog$FLAG)
                
# Convert Missing Values to NA for non-dates
bog$WTE <- ifelse((trimws(as.character(bog$WTE))==trimws("NA")),NA,bog$WTE)  

# Select just the S2 watershed
bog = bog %>%
  filter(PEATLAND == 'S2')

# import temperature data
tempURL  <- "https://pasta.lternet.edu/package/data/eml/edi/583/4/982d194dd32adb4419b14e936056c26c" 
tempInfile <- tempfile()
try(download.file(tempURL,tempInfile,method="curl"))
if (is.na(file.size(tempInfile))) download.file(tempURL,tempInfile,method="auto")

temp <-read.csv(tempInfile, header=F, skip=1, sep=",",
                col.names=c("Date", "STATION", "MAXC",  "MINC",  "FLAG"), check.names=TRUE)
               
unlink(tempInfile)
tmp1Date<-as.Date(temp$Date,format="%Y-%m-%d")  

head(temp)


#Snow depths
snow = read.csv('./mef-data/daily/GR_ForestryLab_snow_daily.csv', 
                na.strings = c('M', 'T'))
snow$Date<-as.POSIXct(snow$Date,format="%Y-%m-%d")

```

## Snow Year Aggregation

Summarized the amount of snow that fell between October 1st and September 30th of the following year and allocates a 'high snow' and 'low snow' depending on if the total snowfall input is above or below the 70 year mean. 

```{r}
#Aggregate to snow year and sum snow inputs
snowAnnual = snow %>%
  mutate(snowYear = ifelse(month(Date) < 10, year(Date), year(Date) + 1)) %>%
  filter(snowYear > 1948) %>%
  group_by(snowYear) %>%
  summarise(totSnow = sum(Snow..inches., na.rm=TRUE))

snowAnnual = snowAnnual %>%
  mutate(typeS = ifelse(totSnow > mean(snowAnnual$totSnow), 'High', 'Low')) %>%
  mutate(year = snowYear) 

head(snowAnnual)
```

Finds the first day of the calendar year in the spring (aka water year in the spring) where snow is persistent on the ground and the last day of the water year where snow is persistent on the year and divides them into high and low snow years so that the 'snow season' can be specified based on a dynamic day where snow disapears, not a fixed date. 

```{r}
#Find first and last non-zero snow and streamflow values
snow_bounds = snow %>%
  mutate(year = year(Date)) %>%
  group_by(year) %>%
  summarise(firstNZ = yday(Date)[which((Snow.Depth..inches. != 0) & !is.na(Snow.Depth..inches.))[1]],
            lastNZ = yday(Date)[tail(which((Snow.Depth..inches. != 0) & (month(Date) < 10) & !is.na(Snow.Depth..inches.)))[5]]) 

snow_bounds_high = snow_bounds %>%
  filter(year %in% snowAnnual$snowYear[snowAnnual$typeS == 'High'])
snow_bounds_low = snow_bounds %>%
  filter(year %in% snowAnnual$snowYear[snowAnnual$typeS == 'Low'])
snow_bounds = rbind(snow_bounds_high, snow_bounds_low)

#stream_bounds = stream %>%
#  mutate(year = year(DateTime)) %>%
#  group_by(year) %>%
  #A threshold is set to eliminate the teeny tiny flows that continue dripping into January
#  summarise(firstNZ = yday(DateTime)[which((q.interval > 0.01) & !is.na(q.interval))[1]],
#            lastNZ = yday(DateTime)[tail(which((q.interval > 0.01) & !is.na(q.interval)))[5]]) 

#stream_bounds_high = stream_bounds %>%
#  filter(year %in% snowAnnual$snowYear[snowAnnual$type == 'High'])
#stream_bounds_low = stream_bounds %>%
#  filter(year %in% snowAnnual$snowYear[snowAnnual$type == 'Low'])
```

#Winter Temperature Aggregation

Similar to the snow dynamics, this allocates the seasonality based on the dynamic snow dates listed above and then takes the mean temperature over the designated 'snow' season. 

```{r}
#Divide seasonal temperatures based on the values in 'snow_bounds'
temp_seasons = temp %>%
  filter(STATION == 'South') %>%
  mutate(MEANC = (MAXC + MINC)/2) %>%
  mutate(year = year(Date)) %>%
  merge(snow_bounds, by = 'year') %>%
  mutate(snowYear = ifelse(month(Date) < 10, year(Date), year(Date) + 1)) %>%
  filter(snowYear > 1948) %>%
  mutate(season = ifelse(yday(Date) > lastNZ & yday(Date) < 183, 'Melt', ifelse(yday(Date) > 182 & yday(Date) < 275, 'Growing', 'Snow'))) %>%
  group_by(season, snowYear) %>%
  summarise(MEANC = mean(MEANC, na.rm = TRUE)) %>%
  rename('year' = 'snowYear')

#Find average snow temperatures
winterTemp = temp_seasons %>%
  filter(season == 'Snow') %>%
  mutate(typeT_snow = ifelse(MEANC > mean(MEANC), 'High', 'Low')) %>%
  rename(winterTempC = MEANC)

#Find average growing temperatures 
growingTemp = temp_seasons %>%
  filter(season == 'Growing') %>%
  mutate(typeT_growing = ifelse(MEANC > mean(MEANC), 'High', 'Low')) %>%
  rename(growingTempC = MEANC)

#Find average melt temperatures 
meltTemp = temp_seasons %>%
  filter(season == 'Melt') %>%
  mutate(typeT_melt = ifelse(MEANC > mean(MEANC), 'High', 'Low')) %>%
  rename(meltTempC = MEANC)
```

```{r}
#Repeat for precipitation
#Divide seasonal precipitation based on the values in 'snow_bounds'
precip_seasons = precip %>%
  mutate(year = year(DATE)) %>%
  merge(snow_bounds, by = 'year') %>%
  mutate(snowYear = ifelse(month(DATE) < 10, year(DATE), year(DATE) + 1)) %>%
  filter(snowYear > 1948) %>%
  mutate(season = ifelse(yday(DATE) > lastNZ & yday(DATE) < 183, 'Melt', ifelse(yday(DATE) > 182 & yday(DATE) < 275, 'Growing', 'Snow'))) %>%
  group_by(season, snowYear) %>%
  summarise(totP = sum(South_PCP, na.rm = TRUE)) %>%
  rename('year' = 'snowYear')

#Find average snow temperatures
winterPrecip = precip_seasons %>%
  filter(season == 'Snow') %>%
  mutate(typeP_snow = ifelse(totP > mean(totP), 'High', 'Low')) %>%
  rename(winterTotP = totP)

#Find average growing temperatures 
growingPrecip = precip_seasons %>%
  filter(season == 'Growing') %>%
  mutate(typeP_growing = ifelse(totP > mean(totP), 'High', 'Low')) %>%
  rename(growingTotP = totP)

#Find average melt temperatures 
meltPrecip = precip_seasons %>%
  filter(season == 'Melt') %>%
  mutate(typeP_melt = ifelse(totP > mean(totP), 'High', 'Low')) %>%
  rename(meltTotPrecip = totP)

#Total annual precipitation
annualPrecip = precip_seasons %>%
  group_by(year) %>%
  summarise(totP = sum(totP)) %>%
  mutate(typeP = ifelse(totP > mean(totP), 'High', 'Low'))
  
```

Aggreagate everything to a seasonal basis and merge into one dataframe.

```{r}
#Aggregate to seasonal 
# 92 for April 1st
#105 for April 15th
precip_seasonal = data.frame(precip) %>%
  mutate(year = year(DATE)) %>%
  mutate(snowYear = ifelse(month(DATE) < 10, year(DATE), year(DATE) + 1)) %>%
  merge(snow_bounds, by = 'year') %>%
  mutate(season = ifelse(yday(DATE) > lastNZ & yday(DATE) < 183, 'Melt', ifelse(yday(DATE) > 182 & yday(DATE) < 275, 'Growing', 'Snow'))) %>%
  group_by(season, snowYear) %>%
  summarise(South_PCP = sum(South_PCP))%>%
  rename('year' = 'snowYear')

stream_seasonal = data.frame(stream) %>%
  mutate(year = year(Date)) %>%
  mutate(snowYear = ifelse(month(Date) < 10, year(Date), year(Date) + 1)) %>%
  merge(snow_bounds, by = 'year') %>%
  mutate(season = ifelse(yday(Date) > lastNZ & yday(Date) < 183, 'Melt', ifelse(yday(Date) > 182 & yday(Date) < 275, 'Growing', 'Snow'))) %>%
  group_by(season, snowYear) %>%
  summarise(mean_cmDay = mean(Flow..cm.day., na.rm = TRUE)/10) %>%
  rename('year' = 'snowYear')

bog_seasonal = data.frame(bog) %>%
  mutate(year = year(DATE)) %>%
  mutate(snowYear = ifelse(month(DATE) < 10, year(DATE), year(DATE) + 1)) %>%
  merge(snow_bounds, by = 'year') %>%
  mutate(season = ifelse(yday(DATE) > lastNZ & yday(DATE) < 183, 'Melt', ifelse(yday(DATE) > 182 & yday(DATE) < 275, 'Growing', 'Snow'))) %>%
  group_by(season, snowYear) %>%
  summarise(WTE = mean(WTE)) %>%
  rename('year' = 'snowYear')

snow_seasonal = data.frame(snow) %>%
  mutate(year = year(Date)) %>%
  mutate(snowYear = ifelse(month(Date) < 10, year(Date), year(Date) + 1)) %>%
  merge(snow_bounds, by = 'year') %>%
  mutate(season = ifelse(yday(Date) > lastNZ & yday(Date) < 183, 'Melt', ifelse(yday(Date) > 182 & yday(Date) < 275, 'Growing', 'Snow'))) %>%
  group_by(season, snowYear) %>%
  summarise(Snow_in = sum(Snow..inches., na.rm = TRUE)) %>%
  rename('year' = 'snowYear')

seasm_seasonal = seasm %>%
  mutate(year = year(DATE)) %>%
  merge(snow_bounds, by = 'year') %>%
  mutate(season = ifelse(yday(DATE) > lastNZ & yday(DATE) < 183, 'Melt', ifelse(yday(DATE) > 182 & yday(DATE) < 289, 'Growing', 'Snow'))) %>%
  group_by(season, year, LOCATION) %>%
  summarise(PAW = mean(PAW, na.rm = TRUE)) %>%
  select(c(year, LOCATION, season, PAW)) %>%
  pivot_wider(names_from = LOCATION, values_from = c(PAW)) 

#merge into dataset
dat = merge(x = precip_seasonal, stream_seasonal, by = c('season', 'year'), how = 'inner')
dat = merge(dat, bog_seasonal, by = c('season', 'year'), how = 'inner')
dat = merge(dat, snow_seasonal, by = c('season', 'year'), how = 'inner')
dat = merge(dat, temp_seasons, by = c('season', 'year'), how = 'inner')
dat = merge(dat, seasm_seasonal, by = c('season', 'year'), how = 'inner') %>% 
  arrange(year, season) 
```
```{r}
#Clip to 1969 onward (for Nans)
dat_clipped = dat %>%
  filter(year > 1968)

summary(dat_clipped)
```


```{r}
write.csv(dat_clipped, './mef-data/data_compiled_gapfilled_seasonal.csv', row.names = FALSE)
dat_clipped
```

