---
title: "Transfer Entrop Flow Analysis"
author: "Mariel Jones"
date: "2024-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
fig_save = "./Figures/"
load_path = "./mef-data/"
```

```{r}
library(astsa)
library(dplyr)
library(xts)
library(lubridate)
library(RTransferEntropy)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(forecastML)
library(cowplot)
library(reshape2)
library(zoo)
library(infotheo)
library(stringr)
library(latex2exp)
```

### Data Import from EDI

```{r, echo = False}
## Precipitation
precipURL  <- "https://pasta.lternet.edu/package/data/eml/edi/849/5/4f7e0b3029833c26d64fda3d23037d1d" 
precipInfile <- tempfile()
try(download.file(precipURL,precipInfile,method="curl"))
if (is.na(file.size(precipInfile))) download.file(precipURL,precipInfile,method="auto")

precip <-read.csv(precipInfile , header=F, skip=1, sep = ",", 
               col.names=c("TIMESTAMP", "South_PCP", "South_Flag"),check.names=TRUE)
               
unlink(precipInfile)
tmp3TIMESTAMP<-as.POSIXct(precip$TIMESTAMP,format="%Y-%m-%d %H:%M:%S")
# Keep the new dates only if they all converted correctly
if(nrow(precip[precip$TIMESTAMP != "",]) == length(tmp3TIMESTAMP[!is.na(tmp3TIMESTAMP)])){precip$TIMESTAMP <- tmp3TIMESTAMP } else {print("Date conversion failed for precip$TIMESTAMP. Please inspect the data and do the date conversion yourself.")}                                 
if (class(precip$South_PCP)=="factor") precip$South_PCP <-as.numeric(levels(precip$South_PCP))[as.integer(precip$South_PCP) ]               
if (class(precip$South_PCP)=="character") precip$South_PCP <-as.numeric(precip$South_PCP)
if (class(precip$South_Flag)!="factor") precip$South_Flag<- as.factor(precip$South_Flag)
                
# Convert Missing Values to NA for non-dates
precip$South_PCP <- ifelse((trimws(as.character(precip$South_PCP))==trimws("NA")),NA,precip$South_PCP)         
## 10 minute soil moisture
smURL  <- "https://pasta.lternet.edu/package/data/eml/edi/989/3/e25d4f15276a3158af412758502a59b0" 
smInfile <- tempfile()
try(download.file(smURL,smInfile,method="curl"))
if (is.na(file.size(smInfile))) download.file(smURL,smInfile,method="auto")

sm <-read.csv(smInfile, header=F, skip=1, sep=",", 
              col.names=c("TIMESTAMP", "S2S_UP_SH", "S2S_UP_DP", "S2S_MI_SH", 
                          "S2S_MI_DP", "S2S_LO_SH", "S2S_LO_DP", "S2N_UP_SH",     
                          "S2N_UP_DP", "S2N_MI_SH", "S2N_MI_DP", "S2N_LO_SH",     
                          "S2N_LO_DP"), check.names=TRUE)
               
unlink(smInfile)
tmp2TIMESTAMP<-as.POSIXct(sm$TIMESTAMP,format="%Y-%m-%d %H:%M:%S")
# Keep the new dates only if they all converted correctly
if(nrow(sm[sm$TIMESTAMP != "",]) == length(tmp2TIMESTAMP[!is.na(tmp2TIMESTAMP)])){sm$TIMESTAMP <- tmp2TIMESTAMP } else {print("Date conversion failed for sm$TIMESTAMP. Please inspect the data and do the date conversion yourself.")}    

                                
## Bog well WTE
bogURL  <- "https://pasta.lternet.edu/package/data/eml/edi/562/3/671f15337a677da71852de506a8d9b05" 
bogInfile <- tempfile()
try(download.file(bogURL,bogInfile,method="curl"))
if (is.na(file.size(bogInfile))) download.file(bogURL,bogInfile,method="auto")

bog <-read.csv(bogInfile, header=F,skip=1, sep=",",
               col.names=c("PEATLAND", "DATE", "WTE", "FLAG"), check.names=TRUE)
               
unlink(bogInfile)
if (class(bog$PEATLAND)!="factor") bog$PEATLAND<- as.factor(bog$PEATLAND)                                   
tmp1DATE<-as.Date(bog$DATE,format="%Y-%m-%d")
# Keep the new dates only if they all converted correctly
if(nrow(bog[bog$DATE != "",]) == length(tmp1DATE[!is.na(tmp1DATE)])){bog$DATE <- tmp1DATE } else {print("Date conversion failed for bog$DATE. Please inspect the data and do the date conversion yourself.")}
if (class(bog$WTE)=="factor") bog$WTE <-as.numeric(levels(bog$WTE))[as.integer(bog$WTE) ]               
if (class(bog$WTE)=="character") bog$WTE <-as.numeric(bog$WTE)
if (class(bog$FLAG)!="factor") bog$FLAG<- as.factor(bog$FLAG)
                
# Convert Missing Values to NA for non-dates
bog$WTE <- ifelse((trimws(as.character(bog$WTE))==trimws("NA")),NA,bog$WTE)  

# Select just the S2 watershed
bog = bog %>%
  filter(PEATLAND == 'S2')



## Lagg transition well WTE
laggwellURL  <- "https://pasta.lternet.edu/package/data/eml/edi/1126/1/2532f551e6e5a97aee8f140fb5da8a74" 
laggwellInfile <- tempfile()
try(download.file(laggwellURL,laggwellInfile,method="curl"))
if (is.na(file.size(laggwellInfile))) download.file(laggwellURL,laggwellInfile,method="auto")

laggwell <-read.csv(laggwellInfile, header=F, skip=1, sep=",", quot='"' ,
                    col.names=c("DateTime", "LoggerLevel", "LoggerTemp", "BP",     
                                "Precip", "bogwell", "well_elev", "well_name",     
                                "watershed", "year"), check.names=TRUE)
               
unlink(laggwellInfile)
              
tmp1DateTime<-as.POSIXct(laggwell$DateTime,format="%Y-%m-%d %H:%M:%S" )
# Keep the new dates only if they all converted correctly
if(nrow(laggwell[laggwell$DateTime != "",]) == length(tmp1DateTime[!is.na(tmp1DateTime)])){laggwell$DateTime <- tmp1DateTime } else {print("Date conversion failed for laggwell$DateTime. Please inspect the data and do the date conversion yourself.")}


                                
## Streamflow
streamURL  <- "https://pasta.lternet.edu/package/data/eml/edi/573/1/2aca4b900546e80ed7dd409ff1ad9787" 
streamInfile <- tempfile()
try(download.file(streamURL,streamInfile,method="curl"))
if (is.na(file.size(streamInfile))) download.file(streamURL,streamInfile,method="auto")

stream <-read.csv(streamInfile, header=F, skip=1, sep=",",
                  col.names=c("Peatland", "DateTime", "Stage.ft", "Q.cfs", "q.mmh", "q.interval"),
                  check.names=TRUE)
               
unlink(streamInfile)
if (class(stream$Peatland)!="factor") stream$Peatland<- as.factor(stream$Peatland)                       
tmp1DateTime<-as.POSIXct(stream$DateTime,format="%Y-%m-%d %H:%M:%S")
# Keep the new dates only if they all converted correctly
if(nrow(stream[stream$DateTime != "",]) == length(tmp1DateTime[!is.na(tmp1DateTime)])){stream$DateTime <- tmp1DateTime } else {print("Date conversion failed for stream$DateTime. Please inspect the data and do the date conversion yourself.")} 


## General Met Station Data
metURL  <- "https://pasta.lternet.edu/package/data/eml/edi/859/4/e1e7792f8465017e133535bccbba4b6c" 
metInfile <- tempfile()
try(download.file(metURL,metInfile,method="curl"))
if (is.na(file.size(metInfile))) download.file(metURL,metInfile,method="auto")
met <-read.csv(metInfile, header=F, skip=1, sep=",", 
               col.names=c("TIMESTAMP", "Air_TempC_Avg", "RH",     
                    "Soil_TempC_Avg", "WS_Tot", "WindDir_D",     
                    "WindDir_SD", "PAR_Den_Avg", "Soil_VWC_Avg"),
               check.names=TRUE)
unlink(metInfile)                      
tmp1DateTime<-as.POSIXct(met$TIMESTAMP,format="%Y-%m-%d %H:%M:%S")
# Keep the new dates only if they all converted correctly
if(nrow(met[met$TIMESTAMP != "",]) == length(tmp1DateTime[!is.na(tmp1DateTime)])){met$TIMESTAMP <- tmp1DateTime } else {print("Date conversion failed for stream$DateTime. Please inspect the data and do the date conversion yourself.")} 

## Air Temp
tempURL  <- "https://pasta.lternet.edu/package/data/eml/edi/583/4/982d194dd32adb4419b14e936056c26c" 
tempInfile <- tempfile()
try(download.file(tempURL,tempInfile,method="curl"))
if (is.na(file.size(tempInfile))) download.file(tempURL,tempInfile,method="auto")
temp <-read.csv(tempInfile, header=F, skip=1, sep=",",  
                col.names=c("Date", "STATION", "MAXC",
                            "MINC", "FLAG"),
                check.names=TRUE)
unlink(tempInfile)
tmp1Date<-as.Date(temp$Date,format="%Y-%m-%d")
# Keep the new dates only if they all converted correctly
if(nrow(temp[temp$Date != "",]) == length(tmp1Date[!is.na(tmp1Date)])){temp$Date <- tmp1Date } else {print("Date conversion failed for temp$Date. Please inspect the data and do the date conversion yourself.")}                                                                    
```

Data is then aggregated to daily timescales

```{r}
#Data sets: stream (1962-2017), sm (2008-2023), precip (2011-2024), laggwell (2018-2020), bog (1961-2023), met (2008-2024), temp (1961-2024)

#Resample precip to daily timestep
#round dates down to week
precip$day <- floor_date(precip$TIMESTAMP, "day")

precip_daily = data.frame(precip %>%
  group_by(day) %>%
  summarize(South_PCP = sum(South_PCP)))
```

```{r}
#Resample well data to daily timestep
#round dates down to week
laggwell$day <- floor_date(laggwell$DateTime, "day")

laggwell_daily = data.frame(laggwell) %>%
  group_by(day, well_name) %>%
  summarize(well_elev = mean(well_elev))
```

```{r}
#Resample temperature to daily timestep
#round dates down to week
met$day <- floor_date(met$TIMESTAMP, "day")

met_daily = data.frame(met) %>%
  group_by(day) %>%
  summarise_all(mean)
```

```{r}
#Resample temperature to daily timestep
#round dates down to week
temp$day <- floor_date(temp$Date, "day")

temp_daily = data.frame(temp) %>%
  filter(STATION == 'S2Bog') %>%
  mutate(MEANC = 0.5*(MAXC + MINC))
```

```{r}
#Melt the lagg data so there is a column per well
laggwell_melt = laggwell_daily %>%
  pivot_wider(names_from = 'well_name', values_from = 'well_elev')
```

```{r}
#Resample soil moisture to daily timestep
#round dates down to week
sm$day <- floor_date(sm$TIMESTAMP, "day")

sm_daily = data.frame(sm) %>%
  group_by(day) %>%
  summarise_all(mean) %>%
  select(-TIMESTAMP)
```

```{r}
#Resample streamflow to daily timestep
#round dates down to week
stream$day <- floor_date(stream$DateTime, "day")

stream_daily = data.frame(stream) %>%
  group_by(day) %>%
  summarize(qInterval = mean(q.interval)/10)
```

Merge data into one data frame for IT analysis

```{r}
#Merge into a dataframe by date and clip to full series
dat = merge(x = precip_daily, y = bog, by.x = "day", by.y = "DATE", all = TRUE)
dat = merge(dat, laggwell_melt, by = 'day', all = TRUE)
dat = merge(dat, sm_daily, by = 'day', how = 'outer', all = TRUE)
dat = merge(dat, stream_daily, by = 'day', how = 'outer', all = TRUE)
dat = merge(dat, met_daily, by = 'day', how = 'outer', all = TRUE)
dat = merge(dat, temp_daily, by = 'day', how = 'outer', all = TRUE)

head(dat)
```

We are now ready for gap filling and data analysis.

## Initial Plotting

```{r}
#Initial timeseries plots

#Precip
ggplot(data = dat, aes(x = day, y = South_PCP)) +
  geom_point(color = 'gray') +
  geom_line() +
  xlab('Date') + 
  ylab('Precipitation [cm]') + 
  theme(aspect.ratio = 0.3, 
        panel.background = element_blank())

#Bog WTE
ggplot(data = dat, aes(x = day, y = WTE)) + 
  geom_point(color = 'gray') + 
  geom_line() + 
  xlab('Date') + 
  ylab('Bog Water Table Elevation [m]') +
  theme(aspect.ratio = 0.3, 
        panel.background = element_blank())

#LAGG WTE
ggplot(data = dat) +
  geom_line(aes(x = day, y = KF42W), color = 'darkolivegreen1') + 
  geom_line(aes(x = day, y = KF43W), color = 'cadetblue1') + 
  geom_line(aes(x = day, y = KF45W), color = 'brown1') +
  geom_line(aes(x = day, y = S2S1), color = 'darkolivegreen') + 
  geom_line(aes(x = day, y = S2S2), color = 'cadetblue3') + 
  geom_line(aes(x = day, y = S2S3), color = 'brown') +
  xlab('Date') +
  ylab('Lagg Water Table Elevation [m]') + 
  theme(legend.position = 'none', 
        aspect.ratio = 0.3, 
        panel.background = element_blank())

#SOIL 
ggplot(data = dat) +
  geom_line(aes(x = day, y = S2S_UP_SH), color = 'darkolivegreen1') + 
  geom_line(aes(x = day, y = S2S_UP_DP), color = 'darkolivegreen1', linetype = 2) + 
  geom_line(aes(x = day, y = S2S_MI_SH), color = 'darkolivegreen3') +
  geom_line(aes(x = day, y = S2S_MI_DP), color = 'darkolivegreen3', linetype = 2) + 
  geom_line(aes(x = day, y = S2S_LO_SH), color = 'darkolivegreen') + 
  geom_line(aes(x = day, y = S2S_LO_DP), color = 'darkolivegreen', linetype = 2) +
  geom_line(aes(x = day, y = S2N_UP_SH), color = 'darkslategray1') + 
  geom_line(aes(x = day, y = S2N_UP_DP), color = 'darkslategray1', linetype = 2) + 
  geom_line(aes(x = day, y = S2N_MI_SH), color = 'darkslategray3') +
  geom_line(aes(x = day, y = S2N_MI_DP), color = 'darkslategray3', linetype = 2) + 
  geom_line(aes(x = day, y = S2N_LO_SH), color = 'darkslategray') + 
  geom_line(aes(x = day, y = S2N_LO_DP), color = 'darkslategray', linetype = 2) +
  xlab('Date') +
  ylab('Soil Moisture [m3/m3]') +
  theme(legend.position = 'none', 
        aspect.ratio = 0.3, 
        panel.background = element_blank()) 

#Streamflow
ggplot(data = dat, aes(x = day, y = qInterval)) +
  geom_point(color = 'gray') + 
  geom_line() + 
  xlab('Date') + 
  ylab('Streamflow [cm]') + 
  theme(aspect.ratio = 0.3,
        panel.background = element_blank())

#Air Temperature
ggplot(data = dat, aes(x = day, y = MEANC)) +
  geom_point(color = 'gray') + 
  geom_line() + 
  xlab('Date') + 
  ylab('Mean Daily Temperature [C]') + 
  theme(aspect.ratio = 0.3,
        panel.background = element_blank())

```

## Clip to timescales for IT analysis

Leaving out the lagg water table levels, clip the data to the longest time length possible (~2012 - 2022)

```{r}
data_clipped = dat %>%
  select(-c('KF42W', 'KF43W', 'KF45W', 'S2S1', 'S2S2', 'S2S3', 'S6S1', 'S6S2', 'S6S3', 'S6N1', 'S6N2', 'S6N3'))

head(data_clipped)
```


```{r}
data = merge(data_clipped,
             data.frame(times = floor_date(seq(from = as.POSIXct('2011-02-03'), to = as.POSIXct('2017-02-28'), by = "day"), 'day')), 
             by.x = 'day', by.y = 'times', 
             all.y = TRUE)
```

```{r}

#Precip
ggplot(data = data, aes(x = day, y = South_PCP)) +
  geom_point(color = 'gray') +
  geom_line() +
  xlab('Date') + 
  ylab('Precipitation [cm]') + 
  theme(aspect.ratio = 0.3, 
        panel.background = element_blank())

#Bog WTE
ggplot(data = data, aes(x = day, y = WTE)) + 
  geom_point(color = 'gray') + 
  geom_line() + 
  xlab('Date') + 
  ylab('Bog Water Table Elevation [m]') +
  theme(aspect.ratio = 0.3, 
        panel.background = element_blank())


#SOIL 
ggplot(data = data) +
  geom_line(aes(x = day, y = S2S_UP_SH), color = 'darkolivegreen1') + 
  geom_line(aes(x = day, y = S2S_UP_DP), color = 'darkolivegreen1', linetype = 2) + 
  geom_line(aes(x = day, y = S2S_MI_SH), color = 'darkolivegreen3') +
  geom_line(aes(x = day, y = S2S_MI_DP), color = 'darkolivegreen3', linetype = 2) + 
  geom_line(aes(x = day, y = S2S_LO_SH), color = 'darkolivegreen') + 
  geom_line(aes(x = day, y = S2S_LO_DP), color = 'darkolivegreen', linetype = 2) +
  geom_line(aes(x = day, y = S2N_UP_SH), color = 'darkslategray1') + 
  geom_line(aes(x = day, y = S2N_UP_DP), color = 'darkslategray1', linetype = 2) + 
  geom_line(aes(x = day, y = S2N_MI_SH), color = 'darkslategray3') +
  geom_line(aes(x = day, y = S2N_MI_DP), color = 'darkslategray3', linetype = 2) + 
  geom_line(aes(x = day, y = S2N_LO_SH), color = 'darkslategray') + 
  geom_line(aes(x = day, y = S2N_LO_DP), color = 'darkslategray', linetype = 2) +
  xlab('Date') +
  ylab('Soil Moisture [m3/m3]') +
  theme(legend.position = 'none', 
        aspect.ratio = 0.3, 
        panel.background = element_blank()) 

#Streamflow
ggplot(data = data, aes(x = day, y = qInterval)) +
  geom_point(color = 'gray') + 
  geom_line() + 
  xlab('Date') + 
  ylab('Streamflow [cm]') + 
  theme(aspect.ratio = 0.3,
        panel.background = element_blank())

#Temperature
ggplot(data = data, aes(x = day, y = MEANC)) + 
  geom_point(color = 'gray') + 
  geom_line() + 
  xlab('Date') + 
  ylab('Air Temperature, [C]') + 
  theme(aspect.ratio = 0.3, 
        panel.background = element_blank())
```

## Aggregate to Monthly timescales

```{r}
#round dates down to week
data$month <- floor_date(data$day, "month")

#group by and summarize
data_means = data %>%
   group_by(month) %>%
   select(c(month, South_PCP, WTE, qInterval, MEANC)) %>%
   summarize_all(mean, na.rm = TRUE) %>%
   rename(meanPrecip = South_PCP, meanWTE = WTE, meanQ = qInterval, meanT = MEANC)

data_totals = data %>%
   group_by(month) %>%
   select(c(month, South_PCP, qInterval)) %>%
   summarize_all(sum, na.rm = TRUE) %>%
   rename(totPrecip = South_PCP, totQ = qInterval)

#Currently grouping all soil data into north and south, deep and shallow resevoirs but will come back and separate these out later
data_soils = data %>%
  mutate(S2S_SH = rowMeans(select(data, c(S2S_UP_SH, S2S_MI_SH, S2S_LO_SH)), na.rm = TRUE)) %>%
  mutate(S2S_DP = rowMeans(select(data, c(S2S_UP_DP, S2S_MI_DP, S2S_LO_DP)), na.rm = TRUE)) %>%
  mutate(S2N_SH = rowMeans(select(data, c(S2N_UP_SH, S2N_MI_SH, S2N_LO_SH)), na.rm = TRUE)) %>%
  mutate(S2N_DP = rowMeans(select(data, c(S2N_UP_DP, S2N_MI_DP, S2N_LO_DP)), na.rm = TRUE)) %>%
  group_by(month) %>%
  select(c(month, S2S_SH, S2S_DP, S2N_SH, S2N_DP)) %>%
  summarize_all(mean, na.rm = TRUE)
```

```{r}
#Merge all data together
data_monthly = merge(data_means, data_totals, by = 'month')
data_monthly = merge(data_monthly, data_soils, by = 'month')
summary(data_monthly)
```

## Monthly plots

```{r}
#Precip
p1 = ggplot(data = data_monthly) +
  geom_point(aes(x = month, y = meanPrecip, color = 'mean Precip')) +
  geom_line(aes(x = month, y = meanPrecip, color = 'mean Precip')) +
  geom_point(aes(x = month, y = totPrecip, color = 'total Precip')) +
  geom_line(aes(x = month, y = totPrecip, color = 'total Precip')) +
  geom_point(aes(x = month, y = meanT, color = 'mean Temp')) +
  geom_line(aes(x = month, y = meanT, color = 'mean Temp')) +
  xlab('Date') + 
  ylab('Precipitation [cm]') + 
  scale_color_manual(" ", values = c('gray', 'black', 'darkred')) + 
  theme(legend.position = 'top',
        aspect.ratio = 0.3, 
        panel.background = element_blank())

#Bog WTE
b1 = ggplot(data = data_monthly, aes(x = month, y = meanWTE)) + 
  geom_point(color = 'gray') + 
  geom_line(color = 'gray') + 
  xlab('Date') + 
  ylab('Bog WTE [m]') +
  theme(legend.position = 'top',
        aspect.ratio = 0.3, 
        panel.background = element_blank())


#SOIL 
s1 = ggplot(data = data_monthly) +
  geom_line(aes(x = month, y = S2S_SH, color = 'S2S SH')) +
  geom_line(aes(x = month, y = S2S_DP, color = 'S2S DP'), linetype = 3) + 
  geom_line(aes(x = month, y = S2N_SH, color = 'S2N SH')) +
  geom_line(aes(x = month, y = S2N_DP, color = 'S2N DP'), linetype = 3) + 
  geom_point(aes(x = month, y = S2S_SH, color = 'S2S SH')) +
  geom_point(aes(x = month, y = S2S_DP, color = 'S2S DP')) + 
  geom_point(aes(x = month, y = S2N_SH, color = 'S2N SH')) +
  geom_point(aes(x = month, y = S2N_DP, color = 'S2N DP')) + 
  xlab('Date') +
  ylab('Soil Moisture [m3/m3]') +
  scale_color_manual(" ", values =  c("darkolivegreen3", "darkolivegreen3", "royalblue", "royalblue")) + 
  theme(legend.position = 'top',
        aspect.ratio = 0.3, 
        panel.background = element_blank()) 

#Streamflow
q1 = ggplot(data = data_monthly) +
  geom_point(aes(x = month, y = meanQ, color = 'mean Q')) +
  geom_line(aes(x = month, y = meanQ, color = 'mean Q')) +
  geom_point(aes(x = month, y = totQ, color = 'total Q')) +
  geom_line(aes(x = month, y = totQ, color = 'total Q')) +
  xlab('Date') + 
  ylab('Streamflow [cm]') + 
  scale_color_manual(" ", values = c('gray', 'black')) +
  theme(legend.position = 'top',
        aspect.ratio = 0.3,
        panel.background = element_blank())

timeseries = plot_grid(p1, b1, s1, q1,
          labels = c('A', 'B', 'C', 'D'),
          label_size = 12,
          align="hv")

save_plot(paste0(fig_save, 'monthlydata_timeseries.pdf'), timeseries, base_height = 5)
save_plot(paste0(fig_save, 'monthlydata_timeseries.jpeg'), timeseries, base_height = 5)
timeseries

```

## Gap filling of soil moisture data
Just a linear gap fill for now because the gaps are no longer than 2 to 3 months. For the daily analysis we may need to come up with a better method.

```{r}
data_monthly_nogaps = data_monthly %>%
  mutate(S2S_DP = na.approx(S2S_DP)) %>%
  mutate(S2S_SH = na.approx(S2S_SH)) %>%
  mutate(S2N_DP = na.approx(S2N_DP)) %>%
  mutate(S2N_SH = na.approx(S2N_SH))%>%
  mutate(meanT = na.approx(meanT))

summary(data_monthly_nogaps)
```

```{r}
ggplot(data = data_monthly_nogaps) +
  geom_line(aes(x = month, y = S2S_SH, color = 'S2S SH')) +
  geom_line(aes(x = month, y = S2S_DP, color = 'S2S DP'), linetype = 3) + 
  geom_line(aes(x = month, y = S2N_SH, color = 'S2N SH')) +
  geom_line(aes(x = month, y = S2N_DP, color = 'S2N DP'), linetype = 3) + 
  geom_point(aes(x = month, y = S2S_SH, color = 'S2S SH')) +
  geom_point(aes(x = month, y = S2S_DP, color = 'S2S DP')) + 
  geom_point(aes(x = month, y = S2N_SH, color = 'S2N SH')) +
  geom_point(aes(x = month, y = S2N_DP, color = 'S2N DP')) + 
  xlab('Date') +
  ylab('Soil Moisture [m3/m3]') +
  scale_color_manual(" ", values =  c("darkolivegreen3", "darkolivegreen3", "royalblue", "royalblue")) + 
  theme(legend.position = 'top',
        aspect.ratio = 0.3, 
        panel.background = element_blank()) 
```

```{r}
ggplot(data = data_monthly_nogaps) +
  geom_point(aes(x = month, y = meanPrecip, color = 'mean Precip')) +
  geom_line(aes(x = month, y = meanPrecip, color = 'mean Precip')) +
  geom_point(aes(x = month, y = totPrecip, color = 'total Precip')) +
  geom_line(aes(x = month, y = totPrecip, color = 'total Precip')) +
  geom_point(aes(x = month, y = meanT, color = 'mean Temp')) +
  geom_line(aes(x = month, y = meanT, color = 'mean Temp')) +
  xlab('Date') + 
  ylab('Precipitation [cm]') + 
  scale_color_manual(" ", values = c('gray', 'darkred', 'black')) + 
  theme(legend.position = 'top',
        aspect.ratio = 0.3, 
        panel.background = element_blank())
```
One final piece of pre-analysis - taking the 5 month anomaly values. This more clearly translates any results that may be minimized or exacerbated by seasonal trends. 

```{r}
data_monthly_anomaly = data_monthly_nogaps %>%
  mutate(across(meanPrecip:S2N_DP, ~ .x - rollmean(.x, k = 5, fill = NA, align = 'left')))


head(data_monthly_nogaps)
head(data_monthly_anomaly)
```

## Plot anomalies

```{r}
#Precip
p1 = ggplot(data = data_monthly_anomaly) +
  geom_point(aes(x = month, y = meanPrecip, color = 'mean Precip')) +
  geom_line(aes(x = month, y = meanPrecip, color = 'mean Precip')) +
  geom_point(aes(x = month, y = totPrecip, color = 'total Precip')) +
  geom_line(aes(x = month, y = totPrecip, color = 'total Precip')) +
  geom_point(aes(x = month, y = meanT, color = 'mean Temp')) +
  geom_line(aes(x = month, y = meanT, color = 'mean Temp')) +
  xlab('Date') + 
  ylab('Precipitation [cm]') + 
  scale_color_manual(" ", values = c('gray', 'black', 'darkred')) + 
  theme(legend.position = 'top',
        aspect.ratio = 0.3, 
        panel.background = element_blank())

#Bog WTE
b1 = ggplot(data = data_monthly_anomaly, aes(x = month, y = meanWTE)) + 
  geom_point(color = 'gray') + 
  geom_line(color = 'gray') + 
  xlab('Date') + 
  ylab('Bog WTE [m]') +
  theme(legend.position = 'top',
        aspect.ratio = 0.3, 
        panel.background = element_blank())


#SOIL 
s1 = ggplot(data = data_monthly_anomaly) +
  geom_line(aes(x = month, y = S2S_SH, color = 'S2S SH')) +
  geom_line(aes(x = month, y = S2S_DP, color = 'S2S DP'), linetype = 3) + 
  geom_line(aes(x = month, y = S2N_SH, color = 'S2N SH')) +
  geom_line(aes(x = month, y = S2N_DP, color = 'S2N DP'), linetype = 3) + 
  geom_point(aes(x = month, y = S2S_SH, color = 'S2S SH')) +
  geom_point(aes(x = month, y = S2S_DP, color = 'S2S DP')) + 
  geom_point(aes(x = month, y = S2N_SH, color = 'S2N SH')) +
  geom_point(aes(x = month, y = S2N_DP, color = 'S2N DP')) + 
  xlab('Date') +
  ylab('Soil Moisture [m3/m3]') +
  scale_color_manual(" ", values =  c("darkolivegreen3", "darkolivegreen3", "royalblue", "royalblue")) + 
  theme(legend.position = 'top',
        aspect.ratio = 0.3, 
        panel.background = element_blank()) 

#Streamflow
q1 = ggplot(data = data_monthly_anomaly) +
  geom_point(aes(x = month, y = meanQ, color = 'mean Q')) +
  geom_line(aes(x = month, y = meanQ, color = 'mean Q')) +
  geom_point(aes(x = month, y = totQ, color = 'total Q')) +
  geom_line(aes(x = month, y = totQ, color = 'total Q')) +
  xlab('Date') + 
  ylab('Streamflow [cm]') + 
  scale_color_manual(" ", values = c('gray', 'black')) +
  theme(legend.position = 'top',
        aspect.ratio = 0.3,
        panel.background = element_blank())

timeseries = plot_grid(p1, b1, s1, q1,
          labels = c('A', 'B', 'C', 'D'),
          label_size = 12,
          align="hv")

save_plot(paste0(fig_save, 'monthlydata_anomaly_timeseries.pdf'), timeseries, base_height = 5)
save_plot(paste0(fig_save, 'monthlydata_anomaly_timeseries.jpeg'), timeseries, base_height = 5)
timeseries
```

## Autocorrelation Functions

```{r}
lags = 24
#Use non-anomalized data for this - data_monthly_nogaps
#Do 12 month seasonal differencing to remove seasonal trends
#Precip
plot(diff(data_monthly_nogaps$totPrecip, 12))
acf2(diff(data_monthly_nogaps$totPrecip, 12), lags)
#Temp
plot(diff(data_monthly_nogaps$meanT, 12))
acf2(diff(data_monthly_nogaps$meanT, 12), lags)
#WTE
plot(diff(data_monthly_nogaps$meanWTE, 12))
acf2(diff(data_monthly_nogaps$meanWTE, 12), lags)
#Soil Moisture
plot(diff(data_monthly_nogaps$S2S_SH, 12))
lines(diff(data_monthly_nogaps$S2S_DP, 12))
lines(diff(data_monthly_nogaps$S2N_SH, 12))
lines(diff(data_monthly_nogaps$S2N_DP, 12))
acf2(diff(data_monthly_nogaps$S2S_SH, 12), lags)
acf2(diff(data_monthly_nogaps$S2S_DP, 12), lags)
acf2(diff(data_monthly_nogaps$S2N_SH, 12), lags)
acf2(diff(data_monthly_nogaps$S2N_DP, 12), lags)
#Streamflow
plot(diff(data_monthly_nogaps$totQ, 12))
acf2(diff(data_monthly_nogaps$totQ, 12), lags)
```




