---
title: "Transfer Entrop Flow Analysis"
author: "Mariel Jones"
date: "2024-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
fig_save = "./Figures/"
load_path = "./mef-data/"
```

```{r}
library(astsa)
library(dplyr)
library(xts)
library(lubridate)
library(RTransferEntropy)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(forecastML)
library(cowplot)
library(reshape2)
library(zoo)
library(infotheo)
library(stringr)
library(latex2exp)
library(Hmisc)
library(plotly)
library(multcompView)
```

# Data Import

Including: 10 minute soil moisture, seasonal soil moisture, precipitation, streamflow, bog well WTE, temperature, and snow depth from GR

```{r}
#Import Soil Moisture and PDSI data
## 10 minute soil moisture
smURL  <- "https://pasta.lternet.edu/package/data/eml/edi/989/3/e25d4f15276a3158af412758502a59b0" 
smInfile <- tempfile()
try(download.file(smURL,smInfile,method="curl"))
if (is.na(file.size(smInfile))) download.file(smURL,smInfile,method="auto")

sm <-read.csv(smInfile, header=F, skip=1, sep=",", 
              col.names=c("TIMESTAMP", "S2S_UP_SH", "S2S_UP_DP", "S2S_MI_SH", 
                          "S2S_MI_DP", "S2S_LO_SH", "S2S_LO_DP", "S2N_UP_SH",     
                          "S2N_UP_DP", "S2N_MI_SH", "S2N_MI_DP", "S2N_LO_SH",     
                          "S2N_LO_DP"), check.names=TRUE)
               
unlink(smInfile)
tmp2TIMESTAMP<-as.POSIXct(sm$TIMESTAMP,format="%Y-%m-%d %H:%M:%S")
# Keep the new dates only if they all converted correctly
if(nrow(sm[sm$TIMESTAMP != "",]) == length(tmp2TIMESTAMP[!is.na(tmp2TIMESTAMP)])){sm$TIMESTAMP <- tmp2TIMESTAMP } else {print("Date conversion failed for sm$TIMESTAMP. Please inspect the data and do the date conversion yourself.")} 
#Aggregate to daily
sm$day <- floor_date(sm$TIMESTAMP, "day")
sm_daily = data.frame(sm) %>%
  group_by(day) %>%
  summarise_all(mean) %>%
  select(-TIMESTAMP)


#Seasonal Soil Moisture (previously processed into plant available water)
seasm <-read.csv('./mef-data/seasonal/mef_seasonal_soil_moisture_processed.csv', header=T, sep=",")
         
# Precipitation
precipURL  <- "https://pasta.lternet.edu/package/data/eml/edi/563/6/f3a58ff544a4ddd3475d265da61bf40e" 
precipInfile <- tempfile()
try(download.file(precipURL,precipInfile,method="curl"))
if (is.na(file.size(precipInfile))) download.file(precipURL,precipInfile,method="auto")

precip <-read.csv(precipInfile , header=F, skip=1, sep = ",", 
               col.names=c("DATE", "NADP_PCP", "South_PCP", "North_PCP",     
                    "NADP_Flag", "South_Flag", "North_Flag") ,check.names=TRUE)
unlink(precipInfile)
tmp3DATE<-as.POSIXct(precip$DATE,format="%Y-%m-%d")
# Keep the new dates only if they all converted correctly
if(nrow(precip[precip$DATE != "",]) == length(tmp3DATE[!is.na(tmp3DATE)])){precip$DATE <- tmp3DATE } else {print("Date conversion failed for precip$TIMESTAMP. Please inspect the data and do the date conversion yourself.")}                                 
if (class(precip$South_PCP)=="factor") precip$South_PCP <-as.numeric(levels(precip$South_PCP))[as.integer(precip$South_PCP) ]               
if (class(precip$South_PCP)=="character") precip$South_PCP <-as.numeric(precip$South_PCP)
if (class(precip$South_Flag)!="factor") precip$South_Flag<- as.factor(precip$South_Flag)
                
# Convert Missing Values to NA for non-dates
precip$South_PCP <- ifelse((trimws(as.character(precip$South_PCP))==trimws("NA")),NA,precip$South_PCP)         

# Streamflow
stream <-read.csv('./mef-data/daily/Streamflow_daily.csv', header=T, sep=",")
stream = stream %>%
  filter(Watershed == 'S2')

# Bog well WTE
bogURL  <- "https://pasta.lternet.edu/package/data/eml/edi/562/3/671f15337a677da71852de506a8d9b05" 
bogInfile <- tempfile()
try(download.file(bogURL,bogInfile,method="curl"))
if (is.na(file.size(bogInfile))) download.file(bogURL,bogInfile,method="auto")

bog <-read.csv(bogInfile, header=F,skip=1, sep=",",
               col.names=c("PEATLAND", "DATE", "WTE", "FLAG"), check.names=TRUE)
               
unlink(bogInfile)
if (class(bog$PEATLAND)!="factor") bog$PEATLAND<- as.factor(bog$PEATLAND)                                   
tmp1DATE<-as.Date(bog$DATE,format="%Y-%m-%d")
# Keep the new dates only if they all converted correctly
if(nrow(bog[bog$DATE != "",]) == length(tmp1DATE[!is.na(tmp1DATE)])){bog$DATE <- tmp1DATE } else {print("Date conversion failed for bog$DATE. Please inspect the data and do the date conversion yourself.")}
if (class(bog$WTE)=="factor") bog$WTE <-as.numeric(levels(bog$WTE))[as.integer(bog$WTE) ]               
if (class(bog$WTE)=="character") bog$WTE <-as.numeric(bog$WTE)
if (class(bog$FLAG)!="factor") bog$FLAG<- as.factor(bog$FLAG)
                
# Convert Missing Values to NA for non-dates
bog$WTE <- ifelse((trimws(as.character(bog$WTE))==trimws("NA")),NA,bog$WTE)  

# Select just the S2 watershed
bog = bog %>%
  filter(PEATLAND == 'S2')

# import temperature data
tempURL  <- "https://pasta.lternet.edu/package/data/eml/edi/583/4/982d194dd32adb4419b14e936056c26c" 
tempInfile <- tempfile()
try(download.file(tempURL,tempInfile,method="curl"))
if (is.na(file.size(tempInfile))) download.file(tempURL,tempInfile,method="auto")

temp <-read.csv(tempInfile, header=F, skip=1, sep=",",
                col.names=c("Date", "STATION", "MAXC",  "MINC",  "FLAG"), check.names=TRUE)
               
unlink(tempInfile)
tmp1Date<-as.Date(temp$Date,format="%Y-%m-%d")  

#Snow depths
snow = read.csv('./mef-data/daily/GR_ForestryLab_snow_daily.csv', 
                na.strings = c('M', 'T'))
snow$Date<-as.POSIXct(snow$Date,format="%Y-%m-%d")

```

```{r}
#Load in prepped data
dat = read.csv('./mef-data/data_compiled_gapfilled_seasonal.csv')
head(dat)
```


## Initial Data Exploration

```{r}
#Bunch o' boxplots
par(mfrow = c(2, 5), 
        mar = c(4, 2, 4, 2))

for(col in colnames(dat[2:9])){
  boxplot(dat[, col] ~ dat$season, xlab = ' ')
  title(col)
}
```

## Seasonal Mutual Information

Transfer entropy is now calculated as lagged mutual information by taking a variable from a single season and comparing it to another variable of another season at a specified lag to determine the amount of 'information' or correlation that a variable from the previous season (or two seasons ago) may have on a variable in the current season. For simplicity the 'sink' variable is defined as streamflow and all other variables are considered to be 'source' variables at lags of 1-2 seasons. 

```{r}
data_seasonal = dat %>%
  drop_na() %>%
  relocate(c("mean_cmDay", "year", "season")) %>%
  #Order seasons so they show up in "Snow", "Melt", "Growing" order
  mutate(season = factor(season, levels = c('Growing', 'Melt', 'Snow')))  %>% 
  group_by(year) %>% 
  filter(all(c('Snow', 'Melt', 'Growing') %in% season))

summary(data_seasonal)
```

Summarized the amount of snow that fell between October 1st and September 30th of the following year and allocates a 'high snow' and 'low snow' depending on if the total snowfall input is above or below the 70 year mean. 

```{r}
#Aggregate to snow year and sum snow inputs
snowAnnual = snow %>%
  mutate(snowYear = ifelse(month(Date) < 10, year(Date), year(Date) + 1)) %>%
  filter(snowYear > 1948) %>%
  group_by(snowYear) %>%
  summarise(totSnow = sum(Snow..inches., na.rm=TRUE))

snowAnnual = snowAnnual %>%
  mutate(typeS = ifelse(totSnow > mean(snowAnnual$totSnow), 'High', 'Low')) %>%
  mutate(year = snowYear) 

#Find first and last non-zero snow and streamflow values
snow_bounds = snow %>%
  mutate(year = year(Date)) %>%
  group_by(year) %>%
  summarise(firstNZ = yday(Date)[which((Snow.Depth..inches. != 0) & !is.na(Snow.Depth..inches.))[1]],
            lastNZ = yday(Date)[tail(which((Snow.Depth..inches. != 0) & (month(Date) < 10) & !is.na(Snow.Depth..inches.)))[5]]) 

snow_bounds_high = snow_bounds %>%
  filter(year %in% snowAnnual$snowYear[snowAnnual$typeS == 'High'])
snow_bounds_low = snow_bounds %>%
  filter(year %in% snowAnnual$snowYear[snowAnnual$typeS == 'Low'])
snow_bounds = rbind(snow_bounds_high, snow_bounds_low)

```

```{r}
#Divide seasonal temperatures based on the values in 'snow_bounds'
temp_seasons = temp %>%
  filter(STATION == 'South') %>%
  mutate(MEANC = (MAXC + MINC)/2) %>%
  mutate(year = year(Date)) %>%
  merge(snow_bounds, by = 'year') %>%
  mutate(snowYear = ifelse(month(Date) < 10, year(Date), year(Date) + 1)) %>%
  filter(snowYear > 1948) %>%
  mutate(season = ifelse(yday(Date) > lastNZ & yday(Date) < 183, 'Melt', ifelse(yday(Date) > 182 & yday(Date) < 275, 'Growing', 'Snow'))) %>%
  group_by(season, snowYear) %>%
  summarise(MEANC = mean(MEANC, na.rm = TRUE)) %>%
  rename('year' = 'snowYear')

#Find average snow temperatures
winterTemp = temp_seasons %>%
  filter(season == 'Snow') %>%
  mutate(typeT_snow = ifelse(MEANC > mean(MEANC), 'High', 'Low')) %>%
  rename(winterTempC = MEANC)

#Find average growing temperatures 
growingTemp = temp_seasons %>%
  filter(season == 'Growing') %>%
  mutate(typeT_growing = ifelse(MEANC > mean(MEANC), 'High', 'Low')) %>%
  rename(growingTempC = MEANC)

#Find average melt temperatures 
meltTemp = temp_seasons %>%
  filter(season == 'Melt') %>%
  mutate(typeT_melt = ifelse(MEANC > mean(MEANC), 'High', 'Low')) %>%
  rename(meltTempC = MEANC)
```

```{r}
#Sort into high and low snow year data frames (based on calendar year, not snow year)
#Snow only
highSnow_seasonal = data_seasonal %>%
  filter(year %in% snowAnnual$snowYear[snowAnnual$typeS == 'High'])

lowSnow_seasonal = data_seasonal %>%
  filter(year %in% snowAnnual$snowYear[snowAnnual$typeS == 'Low'])

#Temp only
highTemp_seasonal = data_seasonal %>%
  filter(year %in% winterTemp$year[winterTemp$typeT_snow == 'High'])

lowTemp_seasonal = data_seasonal %>%
  filter(year %in% winterTemp$year[winterTemp$type_snow == 'Low'])

#Both 
highShighT_seasonal = data_seasonal %>%
  filter((year %in% snowAnnual$snowYear[snowAnnual$typeS == 'High']) &
           (year %in% winterTemp$year[winterTemp$typeT_snow == 'High']))

lowShighT_seasonal = data_seasonal %>%
  filter((year %in% snowAnnual$snowYear[snowAnnual$typeS == 'Low']) &
           (year %in% winterTemp$year[winterTemp$typeT_snow == 'High']))

highSlowT_seasonal = data_seasonal %>%
  filter((year %in% snowAnnual$snowYear[snowAnnual$typeS == 'High']) &
           (year %in% winterTemp$year[winterTemp$typeT_snow == 'Low']))

lowSlowT_seasonal = data_seasonal %>%
  filter((year %in% snowAnnual$snowYear[snowAnnual$typeS == 'Low']) &
           (year %in% winterTemp$year[winterTemp$typeT_snow == 'Low']))

head(data_seasonal)
```


```{r}
## Shift time series
lagTS = function(x, y, l){
  #Shift
  shiftedX = lag(x, n = l)
  shiftedY = lag(y, n = 0)
  
  #Merge into data frame and clip to non-na
  dat = data.frame(shiftedX, shiftedY) %>%
    drop_na()
  
  return(dat)
}

#Seasonal calculation
calc_seasonalQ_contribution_Conditional = function(var, lag, data, subset = FALSE, listYears = NULL){
    t = c()
    
    #Sort
    data = data[order(data$year, data$season),]
    
    #Lag timeseries
    lagDat = lagTS(data[, var], data$mean_cmDay, lag)
    
    if(lag > 0){
      #Lag autocorrelation
      lagY = lagTS(data$mean_cmDay, data$mean_cmDay, lag-1)
          
      #Attach monthly values
      lagDat = cbind(lagDat, season = data$season[0:(length(data$season)-lag)],
                     z = lagY$shiftedY[0:(length(lagY$shiftedY) - 1)], 
                     years = data$year[0:(length(data$year)-lag)])
    }
    else if(lag == 0){
      #lag autocorrelation
      lagY = lagTS(data$mean_cmDay, data$mean_cmDay, 1)
      
      #Attach monthly values
      lagDat = cbind(lagDat[2:length(lagDat$shiftedY),],
                     season = data$season[2:length(data$season)],
                     z = lagY$shiftedX, 
                     years = data$year[2:length(data$year)])
          
    }


    for(i in c('Snow', 'Melt', 'Growing')){
        #filter season
        lagDat_seasonal = lagDat %>% 
          filter(season == i)
        
        #filter years
        if(subset == TRUE){
          lagDat_seasonal = lagDat_seasonal %>%
            filter(years %in% listYears)
        }
        

        #calculate mutual information
        t = append(t, natstobits(condinformation(X = discretize(lagDat_seasonal[, var]),
                                          Y = discretize(lagDat_seasonal$shiftedY), 
                                          S = discretize(lagDat_seasonal$z))))
        
        #print(paste("Variable:", var, ", Lag:", lag, ", Season:", i, ", MI:", temp[length(temp)]))
    }
    
    return(t)
}

calc_seasonalQ_correlations = function(var, lag, data, subset = FALSE, listYears = NULL){
  t = c()
  
  #Sort
  data = data[order(data$year, data$season),]
    
  #print(paste0("Variable: ", var, " Lag: ", lag))
  #print("-----")
  #Lag timeseries
  lagDat = lagTS(data[, var], data$mean_cmDay, lag)
  
  #Attach monthly values
  lagDat = cbind(lagDat, season = data$season[0:(length(data$season)-lag)], 
                 years = data$year[0:(length(data$year)-lag)])
    
  for(i in c('Snow', 'Melt', 'Growing')){
    #filter seasons
    lagDat_seasonal = lagDat %>% filter(season == i) 
    
    #filter years
    if(subset == TRUE){
        lagDat_seasonal = lagDat_seasonal %>%
          filter(years %in% listYears)
    }

    #correlation
    r = cor(x = lagDat_seasonal[,var], y = lagDat_seasonal$shiftedY)
    #print(paste0("Season: ", i, ", R2 = ", r))
    t = append(t, r)

  }
  
  return(t)
}
```


```{r}
#PLOTTING FUNCTIONS
vars = c( "S2S", "S2E", "MEANC",  "South_PCP", "Snow_in", "WTE")

buff = 0.2

plotMI_modified = function(data, title, years = NULL, label = NULL){
  
    #set factor levels
    data = data %>%
      mutate(season = factor(season, levels = c("Snow", "Melt", "Growing")))
      
    for(lag in seq(0, 2)){
      seasonal_MI_data = data.frame(matrix(ncol = 0, nrow = 3))
      seasonal_R_data = data.frame(matrix(ncol = 0, nrow = 3))
      
      for(v in vars){
        
        if(is.null(years)){
          seasonal_MI_data[v] = calc_seasonalQ_contribution_Conditional(v, lag, data)
          seasonal_R_data[v] = calc_seasonalQ_correlations(v, lag, data)
        }
        else{
          seasonal_MI_data[v] = calc_seasonalQ_contribution_Conditional(v, lag,
                                                                        data, subset = TRUE, listYears = years)
          seasonal_R_data[v] = calc_seasonalQ_correlations(v, lag,
                                                           data, subset = TRUE, listYears = years)
        }
        
      }
      
      #Plot stacked time series (plus shifted data)
      x  = c("Snow", "Melt", "Growing")
      if(lag > 0){
        seasons = c(x[(lag+1:(length(x)-lag))], x[1:lag])
      }
      else if(lag == 0){
        seasons = x
      }

      seasonal_MI_data = seasonal_MI_data %>% 
         mutate(season_source = factor(x, levels = c("Snow", "Melt", "Growing"))) %>%
         mutate(season_target = factor(seasons, levels = c("Snow", "Melt", "Growing"))) %>%
         gather(variable, value, vars)
      print(seasonal_MI_data)
      write.csv(seasonal_MI_data, paste0('./MI data/seasonal_mi_data_PAW_', label, '_lag', lag, '.csv'))
      
      seasonal_R_data = seasonal_R_data %>% 
         mutate(season_source = factor(x, levels = c("Snow", "Melt", "Growing"))) %>%
         mutate(season_target = factor(seasons, levels = c("Snow", "Melt", "Growing"))) %>%
         gather(variable, value, vars)
      
      #Change value sign based on R2 value
      #seasonal_MI_data = seasonal_MI_data %>%
      #  mutate(value = ifelse(seasonal_R_data$value < 0, -value, value))
      
      #1 - streamflow, 2 - frost, 3 - melt
      monthlyplot = ggplot(data = seasonal_MI_data, aes(x = season_target, y = value, fill = factor(variable, levels = vars))) +
        #Plot
        geom_col(position = 'dodge') +
        scale_fill_brewer(palette = "BrBG") + 
        xlab(' ') + 
        ylab('Conditional Mutual Information') +
        ylim(0, 1) + 
        theme(legend.position = 'bottom',
              panel.background = element_blank()) +
        ggtitle(paste0(title, ', Lag ', lag))
    
      print(monthlyplot)
      ggsave(paste0('./Figures/Quadrat Plots/', title, '_PAW_Lag', lag, '.pdf'))
  }
}
```

## Plots

Considering first all data together, not separated by annual climatological or hydrological conditions. 

```{r}
#All data
plotMI_modified(data_seasonal, "All Years")
```

# Frost Analysis by Hydrology and Climatology Variables

Import frost depth data from Marcell:

```{r}
#Upland Frost Depths
frostUplandURL  <- "https://pasta.lternet.edu/package/data/eml/edi/571/3/cf159d50f0e80a99b14830103eec9297" 
frostUplandInfile <- tempfile()
try(download.file(frostUplandURL, frostUplandInfile, method="curl"))
if (is.na(file.size(frostUplandInfile))) download.file(frostUplandURL, frostUplandInfile, method="auto")

frostUplandAll <-read.csv(frostUplandInfile, header=F, skip=1, sep=",",
                       col.names=c("Watershed", "SnowcourseID", "Date", "Point", "FrostDEPTH"),
                       check.names=TRUE)
               
unlink(frostUplandInfile)
		    
# attempting to convert dt2$Date dateTime string to R date structure (date or POSIXct)   
tmp2Date<-as.Date(frostUplandAll$Date, format = "%Y-%m-%d")
# Keep the new dates only if they all converted correctly
if(nrow(frostUplandAll[frostUplandAll$Date != "",]) == length(tmp2Date[!is.na(tmp2Date)])){frostUplandAll$Date <- tmp2Date } else {print("Date conversion failed for dt2$Date. Please inspect the data and do the date conversion yourself.")}   

#Filter only S2
frostUpland = frostUplandAll %>% 
  filter(frostUplandAll$SnowcourseID %in% c('S21', 'S22', 'S23'))
  

#Bog Frost Depths
frostBogURL  <- "https://pasta.lternet.edu/package/data/eml/edi/572/1/fbbdd3f237bbde445450d897172ba556" 
frostBogInfile <- tempfile()
try(download.file(frostBogURL, frostBogInfile, method="curl"))
if (is.na(file.size(frostBogInfile))) download.file(frostBogURL, frostBogInfile,method="auto")

frostBog <-read.csv(frostBogInfile, header=F, skip=1, sep=",", 
                    col.names=c("peatland", "year", "frostDate", "depthToFrost", "frostThickness", "frostPercent"), 
                    check.names=TRUE)
               
unlink(frostBogInfile)
		    
# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
if (class(frostBog$peatland)!="factor") frostBog$peatland<- as.factor(frostBog$peatland)                                   
# attempting to convert dt1$frostDate dateTime string to R date structure (date or POSIXct)                                
tmp1frostDate<-as.Date(frostBog$frostDate, format="%Y-%m-%d")
# Keep the new dates only if they all converted correctly
if(nrow(frostBog[frostBog$frostDate != "",]) == length(tmp1frostDate[!is.na(tmp1frostDate)])){frostBog$frostDate <- tmp1frostDate } else {print("Date conversion failed for dt1$frostDate. Please inspect the data and do the date conversion yourself.")}  

#Filter only S2
frostBog = frostBog %>%
  filter(peatland == 'S2')
```

```{r}
#Annual averages
frostUpland_annual = frostUpland %>%
  mutate(year = year(Date)) %>%
  group_by(year) %>%
  summarise(meanUpFrostDepth = mean(FrostDEPTH, na.rm = TRUE)) %>%
  mutate(typeFrost = ifelse(meanUpFrostDepth > mean(meanUpFrostDepth), 'High', 'Low'))

frostBog_annual = frostBog %>%
  group_by(year) %>%
  summarise(meanBogFrostDepth = mean(frostThickness, na.rm = TRUE)) %>%
  mutate(typeFrost = ifelse(meanBogFrostDepth > mean(meanBogFrostDepth), 'High', 'Low'))
```

# IT analysis using measured frost breakdown

```{r}
breaks = quantile(frostUpland_annual$meanUpFrostDepth, probs = c(0, 1/3, 2/3, 1.0))
frostUpland_annual['frostQuantMeas'] = cut(frostUpland_annual$meanUpFrostDepth, breaks = breaks,
                                        include.lowest = TRUE,
                                        labels = c('Low', 'Med', 'High'))
head(frostUpland_annual)
```

```{r}
#Rerun IT with these designations
print("High Frost")
print("-----------------------------")
plotMI_modified(data_seasonal,'High Frost', frostUpland_annual$year[frostUpland_annual$frostQuantMeas == 'High'], label = 'highfrost')


print("Med Frost")
print("-----------------------------")
plotMI_modified(data_seasonal, 'Med Frost', frostUpland_annual$year[frostUpland_annual$frostQuantMeas == 'Med'], label = 'medfrost')


print("Low Frost")
print("-----------------------------")
plotMI_modified(data_seasonal, 'Low Frost', frostUpland_annual$year[frostUpland_annual$frostQuantMeas == 'Low'], label = 'lowfrost')
```







